---
title: "Data Products and Interactivity with R"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
```{r packs, include = FALSE}
library(tidyverse)
ufo_sightings <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-06-25/ufo_sightings.csv")
#filter to US only
ufo_sightings_us <- ufo_sightings %>% 
  filter(country == 'us')

ufo_colorado <- ufo_sightings_us %>% 
  filter(state == "co")
```

## *What happens once analysis is done?*

## Data products

- Rmd report
- Rmd report with interactive elements
- dashboard, interactive plot
- Shiny app 

## Considerations for end users

- can they interpret output?  
- do they know how to code?  
- do they have R/RStudio installed?  

## Goal  

allow end users to interact with data products without involvement of the developer  

## Today's data  



## Interactivity using R

## html widgets  

- JavaScript libraries for R  
- embed them in Rmd/Shiny  

## plotly  

- interactive ggplot2 

## plotly example  
```{r clean, include = FALSE}
sighting_shape <- ufo_sightings %>% 
  filter(ufo_shape == c("circle", "disk", "fireball"))
```

```{r plotly,echo=TRUE, message = FALSE, warning=FALSE}
library(plotly)
plot <- ggplot(data = sighting_shape, aes(x = ufo_shape)) +
            geom_bar(position = "dodge") +
  labs(title = "UFO Sighting Shape",
        x = "Shape", y = "Count") 
  
ggplotly(plot)
```


## leaflet

- create dynamic maps  
- pan, zoom, annotate  

## leaflet example  

```{r leaflet, message=FALSE, warning=FALSE, echo=FALSE}
library(leaflet)
pal <- colorQuantile("YlOrRd", NULL, n = 8)
leaflet(ufo_colorado) %>% 
  addTiles() %>%
  addCircleMarkers(color = ~pal(encounter_length))
```

## flexdashboard  

- group of visualizations  
- versatile components (static, plotly, leaflet)  

## Making data products meaningful  

- color  
- information  
- size  
- text  
- highlighting  

## Create a portfolio!  




